Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/ neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.}

### 2020-07-14 llncs2e/paper12.tex
corrections

\author{Luciano Melodia\orcidID{0000-0002-7584-7287} \\
\and Richard Lenz\orcidID{0000-0003-1551-4824}}
%
\authorrunning{L. Melodia et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Chair of Computer Science 6\\
Friedrich-Alexander University Erlangen-Nürnberg\\
91058 Erlangen, Deutschland \\
\email{\{luciano.melodia,richard.lenz\}@fau.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present an approach to determine the smallest possible number of neurons in a neural net in such a way that the topology of the input space can be learned sufficiently well. We introduce a general procedure based on persistent homology to investigate topological invariants of the manifold on which we suspect the data set. We specify the required dimensions precisely, assuming that there is a smooth manifold on or near which the data are located. Furthermore, we require that this space is connected and has a commutative group structure in the mathematical sense. These assumptions allow us to derive a decomposition of the underlying space whose topology is well known. We use the representatives of the $k$-dimensional homology groups from the persistence landscape to determine an integer dimension for this decomposition. This number is the dimension of the embedding that is capable of capturing the topology of the data manifold. We derive the theory and validate it experimentally on toy data sets.

@@

\subsubsection*{Differential geometry in neural nets} The manifold of a dense neural net is in most cases Euclidean. Recently there have been results in the direction of other manifolds on which neural nets can operate, such as the spherical neural nets \cite{CohenGKW18}. They operate on a manifold with a commutative group structure.

A metric describes the geometric properties of a manifold. The change of coordinate systems and the learned metric tensor of a smooth manifold during backpropagation was formalized \cite{HauserR17}. Further, it has been shown that dense neural nets cannot approximate all functions arbitrarily precisely \cite{Johnson19}. Thereupon neural nets with residual connections were investigated. Residual nets add the output of one layer to a deeper one, bypassing some of the layers in between. Indeed, they are universal function approximators \cite{LinJ18}. Inspired by finite difference methods, such a residual layer was defined as a forward or backward difference operator on a partition of layers. The state space dimension with residual connections is homeomorphic to $\mathbb{R}^{k \cdot n}$, where $n = \text{dim} \; M$ and $k$ is the number of times the difference operator was used \cite{HauserGJR19}. Such a layer is able to embed in $k\cdot n$ dimensional Euclidean space. The smooth manifold perspective inspired us to investigate its invariants leading into the theory discussed in this paper.

@@

\subsubsection{Simplices} In data analysis, we do not study the topological spaces themselves, but points that we assume are located on or near this space. However, all closed surfaces can be triangulated, i.e. completely covered with simplices. For a proof and illustrations we refer to \cite[p.~102]{hatcher2002algebraic,lume}. The concept of the triangle is too specific, so the \emph{$n$-simplices} are defined as a generalization. They are the smallest convex set in Euclidean space with $(n+1)$-points $v_0,\cdots,v_n$, having no solutions for any system of linear equations in $n$-variables. Thus, we say they lie in \emph{general position with respect to $\mathbb{R}^n$}, because they do not lie on any hyperplane with dimension less than $n$ \cite[p.~103]{hatcher2002algebraic}. We define the $n$-simplex as follows:
\begin{equation}
    \label{simplex}
    [\sigma] = \left\{ [v_0,\cdots,v_n] \in \mathbb{R}^{n+1} \; \bigg\vert \; \sum_{i=0}^{n} v_i = 1 \; \text{and} \; v_i \geq 0 \; \text{for all} \; i \right\}.
\end{equation}
Removing a vertex from $[\sigma]$ results in a $(n-1)$-simplex called \emph{face} of $[\sigma]$. A $0$-simplex is a point, a $1$-simplex is a path between two $0$-simplices, a $2$-simplex is an area enclosed by three $1$-simplices and so on. So they follow the intuition and generalize triangles to polyhedra, including points by definition.

@@

\subsubsection{Associated Abelian groups} Algebraically, the simplices can be realized as a system of linear combinations as $\sum_i \lambda_i [\sigma]_{i}$ following Eq. \ref{simplex}. Together with addition as an operation, they form a group structure called $k$th chain group $\langle C_k, +\rangle$ for the corresponding dimension of the $k$-simplices used. The group is commutative or Abelian. The used coefficients $\lambda_i$ induce an orientation, intuitively a direction of the edges, by negative and positive signs, such that $[\sigma] = -[\tau]$, iff $[\sigma] \sim [\tau]$ and $[\sigma]$ and $[\tau]$ have different orientations. The objects of the chain group are called $k$-chains, for $[\sigma] \in C_k$. The groups are connected by a homomorphism, a mapping which respects the algebraic structure, the boundary operator: $\partial_k: C_k \rightarrow C_{k-1}$, $\partial_k [\sigma] \mapsto \sum_i (-1)^i [v_0,v_1,\cdots,\hat{v}_i, \cdots, v_n]$. The $i$th face is omitted, alternatingly. Using the boundary operator we yield the following sequence of Abelian groups, which is a chain complex cf. Fig. \ref{figure1} a):
\begin{equation}
    0 \xrightarrow[]{\partial_{k+1}} C_k \xrightarrow[]{\partial_{k}} C_{k-1} \xrightarrow[]{\partial_{k-1}} \cdots \xrightarrow[]{\partial_{2}} C_1 \xrightarrow[]{\partial_{1}} C_0 \xrightarrow[]{\partial_{0}} 0.
\end{equation}

@@

The $k$th chain groups contain two more Abelian subgroups that behave normal to it. First, the so-called cycle groups $\langle Z_k, +\rangle$, which are defined as $Z_k := \ker \partial_k = \{[\sigma] \in C_k \; \vert \; \partial_k [\sigma] = \emptyset \}$. This follows the intuition of all elements that form a loop, i.e. have the same start and end point. Some of these loops have the peculiarity of being a boundary of a subcomplex. These elements form the boundary group $\langle B_k, +\rangle$, defined by $B_k := \text{im } \partial_{k+1} = \{[\sigma] \in C_k \; \vert \; \exists [\tau] \in C_{k+1}: [\sigma] = \partial_{k+1}[\tau]\}$. Now we define the homology groups as quotient of groups. Homology -- analogous to homotopy theory -- gives information about connected components, loops and higher dimensional holes in the simplicial complex:
\begin{equation}
    H_k(K) := \frac{\ker \partial_k C_k(K)}{\text{im } \partial_{k+1} C_{k+1}(K)} = \frac{Z_k(K)}{B_k(K)}.
\end{equation}

@@

In practice neural nets are used with a different number of neurons per layer. However, a layer can represent the manifold of the previous one. Manifolds can always be immersed and submersed as long as the rank of the Jacobian of the respective map does not change. Thus, the dimension of the data manifold is equal to the width of the smallest possible layer of a neural net \cite{HauserR17}.

@@

Using Künneth's theorem, we know that the $k$-th singular homology group of a topological product space is isomorphic to the direct sum of the tensor product of $k$th homology groups from its factors, for a proof we refer to \cite[p.~268]{hatcher2002algebraic}:
\begin{align}
    H_k(X \times Y) \simeq \bigoplus_{i+j=k}H_i(X) \otimes H_j(Y).
\end{align}
This applies to all fields, since modules over a field are always free. If we apply the theorem to the decomposition, we get for the space we are looking for
\begin{align}
    \label{homologykunneth}
    &H_k(\mathcal{S}^{1}_{1}\times\cdots\times \mathcal{S}^{1}_{q}\times\mathbb{R}^p) \simeq \\
    \bigoplus_{i_{1} + \cdots + i_{r}=k} &H_{i_{1}}(\mathcal{S}^{1}_{1}) \otimes \cdots \otimes H_{i_{\hat{r}}}(\mathcal{S}^{1}_{q}) \otimes H_{i_{r}}(\mathbb{R}^{p}).
\end{align}
One might wonder to what extent homology groups of a simplicial complex can be used to estimate homology groups of smooth manifolds. Through construction of continuous maps from a simplex into a topological space, so-called singular simplices, the singular homology groups $H_{k}(X)$ are obtained. We refer to Hatcher \cite[p.~102]{hatcher2002algebraic} for a proof. This allows to assign Abelian groups to any topological space. If one constructs smooth mappings instead of continuous ones, one gets smooth singular homology groups $H^{\infty}_{k}(X)$, consider \cite[pp.~473]{lee2013smooth} for a proof. Finally a chain complex can be formed on a smooth manifold over the $p$-forms on $X$ and from this the so-called De Rham cohomology can be defined $H_{\text{dR}}^{k}(X)$, for a proof we refer to \cite[pp.~440]{lee2013smooth}. In summary, we have homology theories for simplicial complexes, topological spaces and on smooth manifolds. Simplicial homology is determined by a simplicial complex, which is merely a rough approximation of the triangulation of the underlying topological space. However, we can draw conclusions about a possible smooth manifold, since the discussed homology groups all have isomorphisms to each other known as the De Rham Theorem, see \cite[pp.~106, pp.~467]{hatcher2002algebraic,lee2013smooth}. By the isomorphism of simplicial, singular, smooth singular and De Rham cohomology, our approach is legitimized.

@@

\section{Experimental setting}
\input{figure2.tex}
We train an auto-encoder using as input heavily noisy images and map them to their noiseless original. We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$, such that the input is weighted $0.5 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$. The layer size is increased for each experiment by $2$, i.e. $2,4,6, \cdots, 784$. $392$ neural nets have been trained $\times 100$. Each line of Fig. \ref{figure2} c), d) represents one neural architecture as averaged loss function.\\[0.4cm]
\textbf{Persistence Landscapes:} We use the Delaunay-complex for the filtration, according to Melodia et al. \cite{lume}. The maximum $\alpha$-square is set to $e^{12}$. The maximal expansion dimension of the simplicial complex is set to $10$. The maximal edge length is set to $0.1$. The persistent landscapes are smoothened by a Gaussian filter $G(x) = 1 / \sqrt{2 \pi \sigma^2} \cdot e^{- x^2 / 2 \sigma^2}$ with $\sigma = 2$ for visualization purposes. We implement the persistence diagrams and persistence landscapes using the \texttt{GUDHI} library \cite{gudhi:urm}. Persistent homology is computed on a \texttt{NVIDIA Quadro P4000}.\\[0.4cm]
\textbf{Neural nets:} We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = 1/n \sum_{i=1}^{n} (\textbf{x}^{(0)}_i-\textbf{y}_i)^2$ as loss function. The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = 1 / 1 + e^{-\mathbf{x}^{(l)}}$ is applied throughout the net to yield the structure of a smooth manifold. Dense neural nets are used with bias term. The datasets are flattened from $(32,32,3)$ into $(3072)$. The first layer is applied to the flattened dataset, which is also adapted by back propagation, i.e. $\varphi^{(0)}(x^{(0)}) =$ $\sigma(\sum_{i=1}^{n} \textbf{W}_{ji}^{(1)} \textbf{x}_{i}^{(0)} + \textbf{b}_{i}^{(1)})$. It transfers the data to the desired embedding dimension. The subsequent layers are implemented as residual invertible layers following Dinh et al. \cite{DinhSB17}. We use invertible architectures to stay on the same differentiable structure during learning. A total of $5$ hidden layers are used, all with the same embedding dimension. The batch size is set to $128$, which should help to better recognize good parametrization. Higher batch size causes explosions of the gradient, which we visualize in Fig. \ref{figure2} c), d). For optimization we use \texttt{ADAM} \cite{KingmaB14} and a learning rate of $10^{-4}$. The dataset is randomly shuffled. We implement in \texttt{Python} $3.7$. We use \texttt{Tensorflow} as backend \cite{tensorflow2015-whitepaper} with \texttt{Keras} as wrapper \cite{chollet2015keras}. The training is conducted on \texttt{Intel Core i}$7$ $9700$\texttt{K} processors.\\[0.4cm]
\textbf{Results:} In Fig. \ref{figure2} c) d) we colored the neural nets according to the chosen embedding, such that the lines show the following dimensions ${\color{red1}\bullet} = [2, \cdots, 148]$, ${\color{red2}\bullet} = [150, \cdots, 198]$, ${\color{red3}\bullet} = [200, \cdots, 292]$ and ${\color{red4}\bullet} = [\dim U \approx 294, \cdots, 784]$.  The embedding dimension can be taken from Tab. \ref{figure3}.\footnote{Due to the construction of invertibility the treshold must be doubled, see \cite{DinhSB17}.} All models below our dimensional treshold $\dim U$ show drastic explosions of the gradient, see Fig. \ref{figure2} c) d). The models above the treshold remain stable. A single net still shows an explosive gradient colored ${\color{red4}\bullet}$ in Fig. \ref{figure2} d). The net showing the deflection of gradient is very close to our threshold, identified as $292$ dimensional embedding. The closer we come to the treshold, the sparser the fluctuations of loss.
\input{figure3.tex}

>>>>

\author{Luciano Melodia\orcidID{0000-0002-7584-7287} \\
\and Richard Lenz\orcidID{0000-0003-1551-4824}}
%
\authorrunning{L. Melodia et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Chair of Computer Science 6\\
Friedrich-Alexander University Erlangen-Nürnberg\\
91058 Erlangen, Deutschland \\
\email{\{luciano.melodia,richard.lenz\}@fau.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper, we present an approach to determine the smallest possible number of neurons in a neural net in such a way that the topology of the input space can be learned sufficiently well. We introduce a general procedure based on persistent homology to investigate topological invariants of the manifold on which we suspect the data set. We specify the required dimensions precisely, assuming that there is a smooth manifold on or near which the data are located. Furthermore, we require that this space is connected and has a commutative group structure in the mathematical sense. These assumptions allow us to derive a decomposition of the underlying space whose topology is well known. We use the representatives of the $k$-dimensional homology groups from the persistence landscape to determine an integer dimension for this decomposition. This number is the dimension of the embedding that is capable of capturing the topology of the data manifold. We derive the theory and validate it experimentally on toy data sets.

@@

\subsubsection*{Differential geometry in neural nets} The manifold of a dense neural net is in most cases Euclidean. Nevertheless, it does often approximate a manifold with different structure. Recently there have been results in the direction of other manifolds on which neural nets can operate and that fit more the nature of data, such as the spherical neural nets \cite{CohenGKW18}. They operate on a model of a manifold with commutative group structure. Manifolds induce a notion of distance, with symmetric properties, also referred to as metric.

A metric describes the geometric properties of a manifold. The change of coordinate systems and the learned metric tensor of a smooth manifold during backpropagation was formalized \cite{HauserR17}, which shows, that we operate often on a model of a Lie Group. Further, it has been shown that dense neural nets cannot approximate all functions arbitrarily precisely \cite{Johnson19}. Thereupon neural nets with residual connections were investigated. Residual nets add the output of one layer to a deeper one, bypassing some of the layers in between. Indeed, they are universal function approximators \cite{LinJ18}. Inspired by finite difference methods, such a residual layer was defined as a forward or backward difference operator on a partition of layers. The state space dimension with residual connections is homeomorphic to $\mathbb{R}^{k \cdot n}$, where $n = \text{dim} \; M$ and $k$ is the number of times the difference operator was used \cite{HauserGJR19}. Such a layer is able to embed in $k\cdot n$ dimensional Euclidean space. The smooth manifold perspective inspired us to investigate its invariants leading into the theory discussed in this paper.

@@

\subsubsection{Simplices} In data analysis, we do not study the topological spaces themselves, but points that we assume are located on or near this space. However, all closed surfaces can be triangulated, i.e. completely covered with simplices. For a proof and illustrations we refer to \cite[p.~102]{hatcher2002algebraic}\cite{lume}. The concept of the triangle is too specific, so the \emph{$n$-simplices} are defined as a generalization. They are the smallest convex set in Euclidean space with $(n+1)$-points $v_0,\cdots,v_n$, having no solutions for any system of linear equations in $n$-variables. Thus, we say they lie in \emph{general position with respect to $\mathbb{R}^n$}, because they do not lie on any hyperplane with dimension less than $n$ \cite[p.~103]{hatcher2002algebraic}. We define the $n$-simplex as follows:
\begin{equation}
    \label{simplex}
    [\sigma] = \left\{ [v_0,\cdots,v_n] \in \mathbb{R}^{n+1} \; \bigg\vert \; \sum_{i=0}^{n} v_i = 1 \; \text{and} \; v_i \geq 0 \; \text{for all} \; i \right\}.
\end{equation}
Removing a vertex from $[\sigma]$ results in a $(n-1)$-simplex called \emph{face} of $[\sigma]$. A $0$-simplex is a point, a $1$-simplex is a path between two $0$-simplices, a $2$-simplex is an area enclosed by three $1$-simplices and so on. So they follow the intuition and generalize triangles to polyhedra, including points by definition.

@@

\subsubsection{Associated Abelian groups} Algebraically, the simplices can be realized as a system of linear combinations as $\sum_i \lambda_i [\sigma]_{i}$ following Eq. \ref{simplex}. Together with addition as an operation, they form a group structure called $k$-th chain group $\langle C_k, +\rangle$ for the corresponding dimension of the $k$-simplices used. The group is commutative or Abelian. The used coefficients $\lambda_i$ induce an orientation, intuitively a direction of the edges, by negative and positive signs, such that $[\sigma] = -[\tau]$, iff $[\sigma] \sim [\tau]$ and $[\sigma]$ and $[\tau]$ have different orientations. The objects of the chain group are called $k$-chains, for $[\sigma] \in C_k$. The groups are connected by a homomorphism, a mapping which respects the algebraic structure, the boundary operator: $\partial_k: C_k \rightarrow C_{k-1}$, $\partial_k [\sigma] \mapsto \sum_i (-1)^i [v_0,v_1,\cdots,\hat{v}_i, \cdots, v_n]$. The $i$th face is omitted, alternatingly. Using the boundary operator we yield the following sequence of Abelian groups, which is a chain complex cf. Fig. \ref{figure1} a):
\begin{equation}
    0 \xrightarrow[]{\partial_{k+1}} C_k \xrightarrow[]{\partial_{k}} C_{k-1} \xrightarrow[]{\partial_{k-1}} \cdots \xrightarrow[]{\partial_{2}} C_1 \xrightarrow[]{\partial_{1}} C_0 \xrightarrow[]{\partial_{0}} 0.
\end{equation}

@@

The $k$-th chain groups contain two more Abelian subgroups that behave normal to it. First, the so-called cycle groups $\langle Z_k, +\rangle$, which are defined as $Z_k := \ker \partial_k = \{[\sigma] \in C_k \; \vert \; \partial_k [\sigma] = \emptyset \}$. This follows the intuition of all elements that form a loop, i.e. have the same start and end point. Some of these loops have the peculiarity of being a boundary of a subcomplex. These elements form the boundary group $\langle B_k, +\rangle$, defined by $B_k := \text{im } \partial_{k+1} = \{[\sigma] \in C_k \; \vert \; \exists [\tau] \in C_{k+1}: [\sigma] = \partial_{k+1}[\tau]\}$. Now we define the homology groups as quotient of groups. Homology -- analogous to homotopy theory -- gives information about connected components, loops and higher dimensional holes in the simplicial complex:
\begin{equation}
    H_k(K) := \frac{\ker \partial_k C_k(K)}{\text{im } \partial_{k+1} C_{k+1}(K)} = \frac{Z_k(K)}{B_k(K)}.
\end{equation}

@@

In practice, neural nets are used with a different number of neurons per layer. However, a layer can represent the manifold of the previous one. Manifolds can always be immersed and submersed as long as the rank of the Jacobian of the respective map does not change. Thus, the dimension of the data manifold is equal to the width of the smallest possible layer of a neural net \cite{HauserR17}.

@@

Using Künneth's theorem, we know that the $k$-th singular homology group of a topological product space is isomorphic to the direct sum of the tensor product of $k$-th homology groups from its factors, for a proof we refer to \cite[p.~268]{hatcher2002algebraic}:
\begin{align}
    H_k(X \times Y) \simeq \bigoplus_{i+j=k}H_i(X) \otimes H_j(Y).
\end{align}
This applies to all fields, since modules over a field are always free. If we apply the theorem to the decomposition, we get for the space we are looking for
\begin{align}
    \label{homologykunneth}
    &H_k(\mathcal{S}^{1}_{1}\times\cdots\times \mathcal{S}^{1}_{q}\times\mathbb{R}^p) \simeq \\
    \bigoplus_{i_{1} + \cdots + i_{r}=k} &H_{i_{1}}(\mathcal{S}^{1}_{1}) \otimes \cdots \otimes H_{i_{\hat{r}}}(\mathcal{S}^{1}_{q}) \otimes H_{i_{r}}(\mathbb{R}^{p}).
\end{align}
One might wonder to what extent homology groups of a simplicial complex can be used to estimate homology groups of smooth manifolds. Through construction of continuous maps from a simplex into a topological space, so-called singular simplices, the singular homology groups $H_{k}(X)$ are obtained. We refer to Hatcher \cite[p.~102]{hatcher2002algebraic} for a proof. This allows to assign Abelian groups to any topological space. If one constructs smooth mappings instead of continuous ones, one gets smooth singular homology groups $H^{\infty}_{k}(X)$, consider \cite[pp.~473]{lee2013smooth} for a proof. Finally a chain complex can be formed on a smooth manifold over the $p$-forms on $X$ and from this the so-called De Rham cohomology can be defined $H_{\text{dR}}^{k}(X)$, for a proof we refer to \cite[pp.~440]{lee2013smooth}. In summary, we have homology theories for simplicial complexes, topological spaces and on smooth manifolds. Simplicial homology is determined by a simplicial complex, which is merely a rough approximation of the triangulation of the underlying topological space. However, we can draw conclusions about a possible smooth manifold, since the discussed homology groups all have isomorphisms to each other known as the De Rham Theorem, see \cite[pp.~106, pp.~467]{hatcher2002algebraic,lee2013smooth}. By the isomorphism of simplicial, singular, smooth singular and De Rham cohomology, our approach is legitimized.

@@

\section{Experimental setting}
\input{figure2.tex}
We train an auto-encoder using as input heavily noisy images and map them to their noiseless original. We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$, such that the input is weighted $0.5 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$. The layer size is increased for each experiment by $2$, i.e. $2,4,6, \cdots, 784$. $392$ neural nets around a hundred times. Each line of Fig. \ref{figure2} c), d) represents one neural architecture as averaged loss function.\\[0.4cm]
\textbf{Persistence Landscapes:} We use the Delaunay-complex for the filtration, according to Melodia et al. \cite{lume}. The maximum $\alpha$-square is set to $e^{12}$. The maximal expansion dimension of the simplicial complex is set to $10$. The maximal edge length is set to $0.1$. The persistent landscapes are smoothened by a Gaussian filter $G(x) = 1 / \sqrt{2 \pi \sigma^2} \cdot e^{- x^2 / 2 \sigma^2}$ with $\sigma = 2$ for visualization purposes. We implement the persistence diagrams and persistence landscapes using the \texttt{GUDHI} library \cite{gudhi:urm}. Persistent homology is computed on a \texttt{NVIDIA Quadro P4000}.\\[0.4cm]
\textbf{Neural nets:} We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = 1/n \sum_{i=1}^{n} (\textbf{x}^{(0)}_i-\textbf{y}_i)^2$ as loss function. The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = 1 / 1 + e^{-\mathbf{x}^{(l)}}$ is applied throughout the net to yield the structure of a smooth manifold. Dense neural nets are used with bias term. The datasets are flattened from $(32,32,3)$ into $(3072)$ for the purpose of fitting into a dense layer. The first layer is applied to the flattened dataset, which is also adapted by back propagation, i.e. $\varphi^{(0)}(x^{(0)}) =$ $\sigma(\sum_{i=1}^{n} \textbf{W}_{ji}^{(1)} \textbf{x}_{i}^{(0)} + \textbf{b}_{i}^{(1)})$. It transfers the data to the desired embedding dimension. The subsequent layers are implemented as residual invertible layers following Dinh et al. \cite{DinhSB17}. We use invertible architectures to stay on the same differentiable structure during learning. A total of $5$ hidden layers are used, all with the same embedding dimension. The batch size is set to $128$, which should help to better recognize good parametrization. Higher batch size causes explosions of the gradient, which we visualize in Fig. \ref{figure2} c), d). For optimization we use \texttt{ADAM} \cite{KingmaB14} and a learning rate of $10^{-4}$. The dataset is randomly shuffled. We implement in \texttt{Python} $3.7$. We use \texttt{Tensorflow v.2.2.0} as backend \cite{tensorflow2015-whitepaper} with \texttt{Keras} as wrapper \cite{chollet2015keras}. The training is conducted on \texttt{Intel Core i}$7$ $9700$\texttt{K} processors.\\[0.4cm]
\textbf{Results:} In Fig. \ref{figure2} c) d) we colored the neural nets according to the chosen embedding, such that the lines show the following dimensions ${\color{red1}\bullet} = [2, \cdots, 148]$, ${\color{red2}\bullet} = [150, \cdots, 198]$, ${\color{red3}\bullet} = [200, \cdots, 292]$ and ${\color{red4}\bullet} = [\dim U \approx 294, \cdots, 784]$.  The embedding dimension can be taken from Tab. \ref{figure3}.\footnote{Due to the construction of invertibility the treshold must be doubled, see \cite{DinhSB17}.} All models below our dimensional treshold $\dim U$ show drastic explosions of the gradient, see Fig. \ref{figure2} c) d). The models above the treshold remain stable. A single net still shows an explosive gradient colored ${\color{red4}\bullet}$ in Fig. \ref{figure2} d). The net showing the deflection of gradient is very close to our threshold, identified as $292$ dimensional embedding. The closer we come to the treshold, the sparser the fluctuations of loss.
\input{figure3.tex}
